---
title: "Scratch work Ecology Publicaton"
author: Odalys Barrientos
output: html_notebook
---

# Location of file
```{r}
getwd()
library(mosaic)
```
# New Marix of data 
our data is only 175 lines long 
```{r}
EcologyBias <- EcologyBias_Daneshgar_Popo_2_[1:175, ]
EcologyBias
```
# Count of Ecosystems (popo def)
```{r}
cat('Number of Journal Entries in Marine: ' , count(EcologyBias$`Marine, Terrestrial, Freshwater, Mixed` == "Marine"), '\n')
cat('Number of Journal Entries in Terrestrial: ' , count(EcologyBias$`Marine, Terrestrial, Freshwater, Mixed` == "Terrestrial"), '\n')
cat('Number of Journal Entries in Freshwater: ' , count(EcologyBias$`Marine, Terrestrial, Freshwater, Mixed` == "Freshwater"), '\n')
cat('Number of Journal Entries in Mixed: ' , count(EcologyBias$`Marine, Terrestrial, Freshwater, Mixed` == "Mixed"), '\n')
```
# count of Regions of where the work was done
```{r}
cat('Number of Journal Enties by Region:' , '\n', '\n')
cat('Number of Journal Entries in East-North Central: ' , count(EcologyBias$`Regions where the work was done  (Popo defines)` == "East-North Central"), '\n')
cat('Number of Journal Entries in Mid-Atlantic: ' , count(EcologyBias$`Regions where the work was done  (Popo defines)` == "Mid-Atlantic"), '\n')
cat('Number of Journal Entries in Mid-South: ' , count(EcologyBias$`Regions where the work was done  (Popo defines)` == "MidSouth"), '\n')
cat('Number of Journal Entries in Mountain: ' , count(EcologyBias$`Regions where the work was done  (Popo defines)` == "Mountain"), '\n')
cat('Number of Journal Entries in New England: ' , count(EcologyBias$`Regions where the work was done  (Popo defines)` == "New England"), '\n')
cat('Number of Journal Entries in North West: ' , count(EcologyBias$`Regions where the work was done  (Popo defines)` == "NW"), '\n')
cat('Number of Journal Entries in Pacific: ' , count(EcologyBias$`Regions where the work was done  (Popo defines)` == "Pacific"), '\n')
cat('Number of Journal Entries in South Atlantic: ' , count(EcologyBias$`Regions where the work was done  (Popo defines)` == "South Atlantic"), '\n')
cat('Number of Journal Entries in West North Central: ' , count(EcologyBias$`Regions where the work was done  (Popo defines)` == "West North Central"), '\n')
cat('Number of Journal Entries in West South Central: ' , count(EcologyBias$`Regions where the work was done  (Popo defines)` == "West South Central"), '\n')
cat('Number of Journal Entries in META ANALYSIS: ' , count(EcologyBias$`Regions where the work was done  (Popo defines)` == "META ANALYSIS"), '\n')
cat('Number of Journal Entries in Multiple: ' , count(EcologyBias$`Regions where the work was done  (Popo defines)` == "Multiple"), '\n')
cat('Number of Journal Entries in Multiple Ranges: ' , count(EcologyBias$`Regions where the work was done  (Popo defines)` == "Multiple Ranges"), '\n')
```
# count of ecosystems based off ped def
```{r}
cat('Number of Journal Enties by Ecosystem:' , '\n', '\n')
cat('Number of Journal Entries in Coastal Ecosystems: ' , count(EcologyBias$`Ecosystem (Ped defines)` == "Coastal Ecosystems"), '\n')
cat('Number of Journal Entries in Coniferous Forestss: ' , count(EcologyBias$`Ecosystem (Ped defines)` == "Coniferous Forests"), '\n')
cat('Number of Journal Entries in Deciduous and Evergreens: ' , count(EcologyBias$`Ecosystem (Ped defines)` == "Deciduous and Evergreen"), '\n')
cat('Number of Journal Entries in Ecosystems of the Deep Ocean: ' , count(EcologyBias$`Ecosystem (Ped defines)` == "Ecosystems of the Deep Ocean"), '\n')
cat('Number of Journal Entries in Inland Aquatic Ecosystems: ' , count(EcologyBias$`Ecosystem (Ped defines)` == "Inland Aquatic Ecosystems"), '\n')
cat('Number of Journal Entries in CLakes and Reservoirs Ecosystems: ' , count(EcologyBias$`Ecosystem (Ped defines)` == "Lakes and Reservoirs Ecosystems"), '\n')
cat('Number of Journal Entries in Managed Terrestrial Ecosystems: ' , count(EcologyBias$`Ecosystem (Ped defines)` == "Managed Terrestrial Ecosystems"), '\n')
cat('Number of Journal Entries in Polar & Alpine Tundra: ' , count(EcologyBias$`Ecosystem (Ped defines)` == "Polar & Alpine Tundra"), '\n')
cat('Number of Journal Entries in Shrublands: ' , count(EcologyBias$`Ecosystem (Ped defines)` == "Shrublands"), '\n')
cat('Number of Journal Entries in Temperate Deciduous Forests: ' , count(EcologyBias$`Ecosystem (Ped defines)` == "Temperate Deciduous Forests"), '\n')
cat('Number of Journal Entries in Tropical Forests: ' , count(EcologyBias$`Ecosystem (Ped defines)` == "Tropical Forests"), '\n')
cat('Number of Journal Entries in CWetland Forests: ' , count(EcologyBias$`Ecosystem (Ped defines)` == "Wetland Forests"), '\n')
cat('Number of Journal Entries in Coniferous and Hardwood: ' , count(EcologyBias$`Ecosystem (Ped defines)` == "Coniferous and Hardwood"), '\n')
cat('Number of Journal Entries in Coral Reefs: ' , count(EcologyBias$`Ecosystem (Ped defines)` == "Coral Reefs"), '\n')
cat('Number of Journal Entries in Ecosystems of the Continental Shelves: ' , count(EcologyBias$`Ecosystem (Ped defines)` == "Ecosystems of the Continental Shelves"), '\n')
cat('Number of Journal Entries in Estuaries and Enclosed Seas: ' , count(EcologyBias$`Ecosystem (Ped defines)` == "Estuaries and Enclosed Seas"), '\n')
cat('Number of Journal Entries in Intertidal and Littoral Ecosystems: ' , count(EcologyBias$`Ecosystem (Ped defines)` == "Intertidal and Littoral Ecosystems"), '\n')
cat('Number of Journal Entries in Managed Aquatic Ecosystems: ' , count(EcologyBias$`Ecosystem (Ped defines)` == "Managed Aquatic Ecosystems"), '\n')
cat('Number of Journal Entries in Natural Grasslands: ' , count(EcologyBias$`Ecosystem (Ped defines)` == "Natural Grasslands"), '\n')
cat('Number of Journal Entries in River and Stream Ecosystem: ' , count(EcologyBias$`Ecosystem (Ped defines)` == "River and Stream Ecosystem"), '\n')
cat('Number of Journal Entries in Temperate Broad Leaved Evergreen: ' , count(EcologyBias$`Ecosystem (Ped defines)` == "Temperate Broad Leaved Evergreen"), '\n')
cat('Number of Journal Entries in Temperate Deserts and Semi Deserts: ' , count(EcologyBias$`Ecosystem (Ped defines)` == "Temperate Deserts and Semi Deserts"), '\n')
cat('Number of Journal Entries in Tropical Savannas: ' , count(EcologyBias$`Ecosystem (Ped defines)` == "CTropical Savannas"), '\n')
cat('Number of Journal Entries in Multiple Ecosystems: ' , count(EcologyBias$`Ecosystem (Ped defines)` == "Multiple"), '\n')
cat('Number of Journal Entries in META-ANALYSIS: ' , count(EcologyBias$`Ecosystem (Ped defines)` == "META-ANALYSIS"), '\n')
```
# count of phylum
```{r}
cat('Number of Journal Enties by Phylum:' , '\n', '\n')
cat('Number of Journal Entries in Annelida: ' , count(EcologyBias$Phylum == "Annelida"), '\n')
cat('Number of Journal Entries in Anthophyta: ' , count(EcologyBias$Phylum == "Anthophyta"), '\n')
cat('Number of Journal Entries in Arthropoda: ' , count(EcologyBias$Phylum == "Arthropoda"), '\n')
cat('Number of Journal Entries in Chordata: ' , count(EcologyBias$Phylum == "Chordata"), '\n')
cat('Number of Journal Entries in Cnidaria: ' , count(EcologyBias$Phylum == "Cnidaria"), '\n')
cat('Number of Journal Entries in Coniferophyta: ' , count(EcologyBias$Phylum == "Coniferophyta"), '\n')
cat('Number of Journal Entries in Echinodermata: ' , count(EcologyBias$Phylum == "Echinodermata"), '\n')
cat('Number of Journal Entries in Magnoliophyta: ' , count(EcologyBias$Phylum == "Magnoliophyta"), '\n')
cat('Number of Journal Entries in Mollusca: ' , count(EcologyBias$Phylum == "Mollusca"), '\n')
cat('Number of Journal Entries in Ochrophyta: ' , count(EcologyBias$Phylum == "Ochrophyta"), '\n')
cat('Number of Journal Entries in Oomycota: ' , count(EcologyBias$Phylum == "Oomycota"), '\n')
cat('Number of Journal Entries in Spermatophyta: ' , count(EcologyBias$Phylum == "Spermatophyta"), '\n')
cat('Number of Journal Entries in Tracheophyta: ' , count(EcologyBias$Phylum == "Tracheophyta"), '\n')
cat('Number of Journal Entries in Multiple: ' , count(EcologyBias$Phylum == "Multiple"), '\n')
```
# matrix of phylum Genus and species
```{r}
classes <- EcologyBias[,18:20]
classes

# Taking away NA of Classes

class.clean <- na.omit(classes)
```
# counts by genus
```{r}
cat('Number of Journal Enties by Genus:' , '\n', '\n')
cat('Number of Journal Entries in Acacia: ' , count(class.clean[,2] == "Acacia"), '\n')
cat('Number of Journal Entries in Acer: ' , count(class.clean[,2] == "Acer"), '\n')
cat('Number of Journal Entries in Andropogon: ' , count(class.clean[,2] == "Andropogon"), '\n')
cat('Number of Journal Entries in Anthocharis: ' , count(class.clean[,2] == "Anthocharis"), '\n')
cat('Number of Journal Entries in Aplexa: ' , count(class.clean[,2]== "Aplexa"), '\n')
cat('Number of Journal Entries in Artemisia: ' , count(class.clean[,2] == "Artemisia"), '\n')
cat('Number of Journal Entries in Azteca: ' , count(class.clean[,2] == "Azteca"), '\n')
cat('Number of Journal Entries in Bombus: ' , count(class.clean[,2] == "Bombus"), '\n')
cat('Number of Journal Entries in Bouteloua: ' , count(class.clean[,2]== "Bouteloua"), '\n')
cat('Number of Journal Entries in Brassica: ' , count(class.clean[,2] == "Brassica"), '\n')
cat('Number of Journal Entries in Bromus: ' , count(class.clean[,2] == "Bromus"), '\n')
cat('Number of Journal Entries in Capreolus: ' , count(class.clean[,2] == "Capreolus"), '\n')
cat('Number of Journal Entries in Cladium: ' , count(class.clean[,2] == "Cladium"), '\n')
cat('Number of Journal Entries in Coenobita: ' , count(class.clean[,2] == "Coenobita"), '\n')
```
This list continues I dont think this is a good way to do this. 

```{r}
Spermatophyta.G<- vector()
Spermatophyta.S <- vector()
for(i in 1:59){
 if(class.clean[i,1]== "Spermatophyta"){
   Spermatophyta.G[i] <- class.clean[i,2]
   Spermatophyta.S[i] <- class.clean[i,3]
 }}
Spermatophyta.G.2 <- Spermatophyta.G[-which(sapply(Spermatophyta.G, is.null))]
Spermatophyta.G.2

Spermatophyta.S.2 <- Spermatophyta.S[-which(sapply(Spermatophyta.S, is.null))]
Spermatophyta.S.2

```
These are now the only ones where we have all 3 classes. There are 59 rows so we lost a lot of data. 
```{r}
mult.G<- vector()
mult.S <- vector()
for(i in 1:59){
 if(class.clean[i,1]== "Multiple"){
   mult.G[i] <- class.clean[i,2]
   mult.S[i] <- class.clean[i,3]
 }}
mult.G.2 <- mult.G[-which(sapply(mult.G, is.null))]
mult.G.2

mult.S.2 <- mult.S[-which(sapply(mult.S, is.null))]
mult.S.2

```

```{r}

Chordata.G<- vector()
Chordata.S <- vector()
for(i in 1:59){
 if(class.clean[i,1]== "Chordata"){
   Chordata.G[i] <- class.clean[i,2]
   Chordata.S[i] <- class.clean[i,3]
 }}
Chordata.G.2 <- Chordata.G[-which(sapply(Chordata.G, is.null))]
Chordata.G.2

Chordata.S.2 <- Chordata.S[-which(sapply(Chordata.S, is.null))]
Chordata.S.2

```

