---
title: "Scratch work Ecology Publicaton"
author: Odalys Barrientos
output: html_notebook
---

# Location of file
```{r}
getwd()
library(mosaic)
```
# New Marix of data 
our data is only 175 lines long 
```{r}
EcologyBias <- EcologyBias_Daneshgar_Popo_2_[1:175, ]
EcologyBias
```
# Count of Ecosystems (popo def)
```{r}
cat('Number of Journal Entries in Marine: ' , count(EcologyBias$`Marine, Terrestrial, Freshwater, Mixed` == "Marine"), '\n')
cat('Number of Journal Entries in Terrestrial: ' , count(EcologyBias$`Marine, Terrestrial, Freshwater, Mixed` == "Terrestrial"), '\n')
cat('Number of Journal Entries in Freshwater: ' , count(EcologyBias$`Marine, Terrestrial, Freshwater, Mixed` == "Freshwater"), '\n')
cat('Number of Journal Entries in Mixed: ' , count(EcologyBias$`Marine, Terrestrial, Freshwater, Mixed` == "Mixed"), '\n')
```
# count of Regions of where the work was done
```{r}
cat('Number of Journal Enties by Region:' , '\n', '\n')
cat('Number of Journal Entries in East-North Central: ' , count(EcologyBias$`Regions where the work was done  (Popo defines)` == "East-North Central"), '\n')
cat('Number of Journal Entries in Mid-Atlantic: ' , count(EcologyBias$`Regions where the work was done  (Popo defines)` == "Mid-Atlantic"), '\n')
cat('Number of Journal Entries in Mid-South: ' , count(EcologyBias$`Regions where the work was done  (Popo defines)` == "MidSouth"), '\n')
cat('Number of Journal Entries in Mountain: ' , count(EcologyBias$`Regions where the work was done  (Popo defines)` == "Mountain"), '\n')
cat('Number of Journal Entries in New England: ' , count(EcologyBias$`Regions where the work was done  (Popo defines)` == "New England"), '\n')
cat('Number of Journal Entries in North West: ' , count(EcologyBias$`Regions where the work was done  (Popo defines)` == "NW"), '\n')
cat('Number of Journal Entries in Pacific: ' , count(EcologyBias$`Regions where the work was done  (Popo defines)` == "Pacific"), '\n')
cat('Number of Journal Entries in South Atlantic: ' , count(EcologyBias$`Regions where the work was done  (Popo defines)` == "South Atlantic"), '\n')
cat('Number of Journal Entries in West North Central: ' , count(EcologyBias$`Regions where the work was done  (Popo defines)` == "West North Central"), '\n')
cat('Number of Journal Entries in West South Central: ' , count(EcologyBias$`Regions where the work was done  (Popo defines)` == "West South Central"), '\n')
cat('Number of Journal Entries in META ANALYSIS: ' , count(EcologyBias$`Regions where the work was done  (Popo defines)` == "META ANALYSIS"), '\n')
cat('Number of Journal Entries in Multiple: ' , count(EcologyBias$`Regions where the work was done  (Popo defines)` == "Multiple"), '\n')
cat('Number of Journal Entries in Multiple Ranges: ' , count(EcologyBias$`Regions where the work was done  (Popo defines)` == "Multiple Ranges"), '\n')
```
# count of ecosystems based off ped def
```{r}
cat('Number of Journal Enties by Ecosystem:' , '\n', '\n')
cat('Number of Journal Entries in Coastal Ecosystems: ' , count(EcologyBias$`Ecosystem (Ped defines)` == "Coastal Ecosystems"), '\n')
cat('Number of Journal Entries in Coniferous Forestss: ' , count(EcologyBias$`Ecosystem (Ped defines)` == "Coniferous Forests"), '\n')
cat('Number of Journal Entries in Deciduous and Evergreens: ' , count(EcologyBias$`Ecosystem (Ped defines)` == "Deciduous and Evergreen"), '\n')
cat('Number of Journal Entries in Ecosystems of the Deep Ocean: ' , count(EcologyBias$`Ecosystem (Ped defines)` == "Ecosystems of the Deep Ocean"), '\n')
cat('Number of Journal Entries in Inland Aquatic Ecosystems: ' , count(EcologyBias$`Ecosystem (Ped defines)` == "Inland Aquatic Ecosystems"), '\n')
cat('Number of Journal Entries in CLakes and Reservoirs Ecosystems: ' , count(EcologyBias$`Ecosystem (Ped defines)` == "Lakes and Reservoirs Ecosystems"), '\n')
cat('Number of Journal Entries in Managed Terrestrial Ecosystems: ' , count(EcologyBias$`Ecosystem (Ped defines)` == "Managed Terrestrial Ecosystems"), '\n')
cat('Number of Journal Entries in Polar & Alpine Tundra: ' , count(EcologyBias$`Ecosystem (Ped defines)` == "Polar & Alpine Tundra"), '\n')
cat('Number of Journal Entries in Shrublands: ' , count(EcologyBias$`Ecosystem (Ped defines)` == "Shrublands"), '\n')
cat('Number of Journal Entries in Temperate Deciduous Forests: ' , count(EcologyBias$`Ecosystem (Ped defines)` == "Temperate Deciduous Forests"), '\n')
cat('Number of Journal Entries in Tropical Forests: ' , count(EcologyBias$`Ecosystem (Ped defines)` == "Tropical Forests"), '\n')
cat('Number of Journal Entries in CWetland Forests: ' , count(EcologyBias$`Ecosystem (Ped defines)` == "Wetland Forests"), '\n')
cat('Number of Journal Entries in Coniferous and Hardwood: ' , count(EcologyBias$`Ecosystem (Ped defines)` == "Coniferous and Hardwood"), '\n')
cat('Number of Journal Entries in Coral Reefs: ' , count(EcologyBias$`Ecosystem (Ped defines)` == "Coral Reefs"), '\n')
cat('Number of Journal Entries in Ecosystems of the Continental Shelves: ' , count(EcologyBias$`Ecosystem (Ped defines)` == "Ecosystems of the Continental Shelves"), '\n')
cat('Number of Journal Entries in Estuaries and Enclosed Seas: ' , count(EcologyBias$`Ecosystem (Ped defines)` == "Estuaries and Enclosed Seas"), '\n')
cat('Number of Journal Entries in Intertidal and Littoral Ecosystems: ' , count(EcologyBias$`Ecosystem (Ped defines)` == "Intertidal and Littoral Ecosystems"), '\n')
cat('Number of Journal Entries in Managed Aquatic Ecosystems: ' , count(EcologyBias$`Ecosystem (Ped defines)` == "Managed Aquatic Ecosystems"), '\n')
cat('Number of Journal Entries in Natural Grasslands: ' , count(EcologyBias$`Ecosystem (Ped defines)` == "Natural Grasslands"), '\n')
cat('Number of Journal Entries in River and Stream Ecosystem: ' , count(EcologyBias$`Ecosystem (Ped defines)` == "River and Stream Ecosystem"), '\n')
cat('Number of Journal Entries in Temperate Broad Leaved Evergreen: ' , count(EcologyBias$`Ecosystem (Ped defines)` == "Temperate Broad Leaved Evergreen"), '\n')
cat('Number of Journal Entries in Temperate Deserts and Semi Deserts: ' , count(EcologyBias$`Ecosystem (Ped defines)` == "Temperate Deserts and Semi Deserts"), '\n')
cat('Number of Journal Entries in Tropical Savannas: ' , count(EcologyBias$`Ecosystem (Ped defines)` == "CTropical Savannas"), '\n')
cat('Number of Journal Entries in Multiple Ecosystems: ' , count(EcologyBias$`Ecosystem (Ped defines)` == "Multiple"), '\n')
cat('Number of Journal Entries in META-ANALYSIS: ' , count(EcologyBias$`Ecosystem (Ped defines)` == "META-ANALYSIS"), '\n')
```
# count of phylum
```{r}
cat('Number of Journal Enties by Phylum:' , '\n', '\n')
cat('Number of Journal Entries in Annelida: ' , count(EcologyBias$Phylum == "Annelida"), '\n')
cat('Number of Journal Entries in Anthophyta: ' , count(EcologyBias$Phylum == "Anthophyta"), '\n')
cat('Number of Journal Entries in Arthropoda: ' , count(EcologyBias$Phylum == "Arthropoda"), '\n')
cat('Number of Journal Entries in Chordata: ' , count(EcologyBias$Phylum == "Chordata"), '\n')
cat('Number of Journal Entries in Cnidaria: ' , count(EcologyBias$Phylum == "Cnidaria"), '\n')
cat('Number of Journal Entries in Coniferophyta: ' , count(EcologyBias$Phylum == "Coniferophyta"), '\n')
cat('Number of Journal Entries in Echinodermata: ' , count(EcologyBias$Phylum == "Echinodermata"), '\n')
cat('Number of Journal Entries in Magnoliophyta: ' , count(EcologyBias$Phylum == "Magnoliophyta"), '\n')
cat('Number of Journal Entries in Mollusca: ' , count(EcologyBias$Phylum == "Mollusca"), '\n')
cat('Number of Journal Entries in Ochrophyta: ' , count(EcologyBias$Phylum == "Ochrophyta"), '\n')
cat('Number of Journal Entries in Oomycota: ' , count(EcologyBias$Phylum == "Oomycota"), '\n')
cat('Number of Journal Entries in Spermatophyta: ' , count(EcologyBias$Phylum == "Spermatophyta"), '\n')
cat('Number of Journal Entries in Tracheophyta: ' , count(EcologyBias$Phylum == "Tracheophyta"), '\n')
cat('Number of Journal Entries in Multiple: ' , count(EcologyBias$Phylum == "Multiple"), '\n')
```
# contingency tables of ORIGINAL country, region, state, and continent
```{r}

tally(EcologyBias$`Country where the work is done`) # needs to be fixed
tally(EcologyBias$`Regions where the work was done  (Popo defines)`) # needs to be fixed
tally(EcologyBias$`State or Municipality where the work was done`) # needs to be fixed
tally(EcologyBias$`Continent where work done`) # needs to be fixed

```
# Continent
```{r}
continent.work <- as.vector(EcologyBias$`Continent where work done`)
remove <- c ("Europe & Oceania", "META ANALYSIS","META ANALYSIS!")
continent.work %in% remove
continent.work <- continent.work [! continent.work %in% remove]
continent.work[43] <- "Oceania"
continent.work[61] <- "Oceania"
continent.work <- c(continent.work[1:97], continent.work[99:168])
continent.work
tally(continent.work)
chisq.test(tally(continent.work))

```
# Chisq test country
```{r}
country.work <- as.vector(EcologyBias$`Country where the work is done`)
remove <- c ("Brazil/Costa Rica/Puerto Rico", "Germany & New Zealand","Lesser Antilles/Guadelupe", "META ANALYSIS", "Multiple???", "US and Mexico", "Trinidad and Tobago")
country.work %in% remove
country.work <- country.work [! country.work %in% remove]
country.work[1:5] <- "USA"
country.work[8:11] <- "USA"
country.work <- c(country.work[1:25], country.work[27:34], country.work[36:77], country.work[79:95], country.work[97:121], country.work[123:144], country.work[146:168])
country.work
tally(country.work)
chisq.test(tally(country.work))

```
# Chisq test region
```{r}
region.work <- as.vector(EcologyBias$`Regions where the work was done  (Popo defines)`)
region.work
remove <- c ("Multiple", "META ANALYSIS", "Multiple Ranges")
region.work %in% remove
region.work <- region.work [! region.work %in% remove]
region.work
region.work <- region.work[!is.na(region.work)]
region.work
tally(region.work)
chisq.test(tally(region.work))

```
# Chisq test state
```{r}
state.work <- as.vector(EcologyBias$`State or Municipality where the work was done`)
state.work
remove <- c ("Multiple States", "Multiple", "META ANALYSIS", "OR/CA", "GA/FL", "NY, MA, CT")
state.work %in% remove
state.work <- state.work [! state.work %in% remove]
state.work
state.work <- state.work[!is.na(state.work)]
state.work
tally(state.work)
chisq.test(tally(state.work))

```

# Chisq test ecosystems popo def
```{r}
eco.po <- as.vector(EcologyBias$`Marine, Terrestrial, Freshwater, Mixed`)
eco.po
remove <- c ("Mixed")
eco.po %in% remove
eco.po <- eco.po [! eco.po %in% remove]
eco.po

tally(eco.po)
chisq.test(tally(eco.po))
```

# Chisq ecosystem ped def
```{r}
eco.ped <- as.vector(EcologyBias$`Ecosystem (Ped defines)`)

remove <- c ("Multiple", "META-ANALYSIS")
eco.ped %in% remove
eco.ped <- eco.ped [! eco.ped %in% remove]

eco.ped <- c(eco.ped[1:50], eco.ped[52:169])
eco.ped
tally(eco.ped)
chisq.test(tally(eco.ped))
```

