---
title: "Dr.Daneshgar Data Analysis"
author: Veronica Marquez
output: html_notebook
---

get file location
```{r}
getwd()
```
run library
```{r}
library(mosaic)
```

```{r}
library(readxl)
EcologyBias_Daneshgar_Popo_2_ <- read_excel("/Users/veronicamarquez/Desktop/Dr.D Ecology Bias SC/MA321-Daneshgar/EcologyBias.Daneshgar.Popo.xlsx")
View(EcologyBias_Daneshgar_Popo_2_)
```

```{r}
EcologyBias <- EcologyBias_Daneshgar_Popo_2_[1:175, ]
```

Looking to see how many of each ecosystem exists in the data set 
```{r}
cat("# of Journals in Marine: ", count(EcologyBias$`Marine, Terrestrial, Freshwater, Mixed` == "Marine"), "\n")
cat("# of Journals in Terrestrial: ", count(EcologyBias$`Marine, Terrestrial, Freshwater, Mixed` == "Terrestrial"), "\n")
cat("# of Journals in Frshwater: ", count(EcologyBias$`Marine, Terrestrial, Freshwater, Mixed` == "Freshwater"), "\n")
cat("# of Journals in Mixed: ", count(EcologyBias$`Marine, Terrestrial, Freshwater, Mixed` == "Mixed"), "\n")
```
Notice here that there is a significant difference in the amount of articles that contain each ecosystem. Is this a glimpse into the "Sexy" type of ecosystem? Should I take it a step closer and analyze the ecosystems based on specif regions? If we take just this information, then we definitely some form of bias between these ecosystem in itself. 

##Location of First Author (Region)
```{r}
cat("# of Journals in East-North Central: ", count(EcologyBias$`Location of first author (region)` == "East-North Central"), "\n")
cat("# of Journals in Mid-Atlantic: ", count(EcologyBias$`Location of first author (region)` == "Mid-Atlantic"), "\n")
cat("# of Journals in Mid-South: ", count(EcologyBias$`Location of first author (region)` == "Mid-South"), "\n")
cat("# of Journals in Mountain: ", count(EcologyBias$`Location of first author (region)` == "Mountain"), "\n")
cat("# of Journals in New England: ", count(EcologyBias$`Location of first author (region)` == "New England"), "\n")
cat("# of Journals in North West: ", count(EcologyBias$`Location of first author (region)` == "North West"), "\n")
cat("# of Journals in Pacific: ", count(EcologyBias$`Location of first author (region)` == "Pacific"), "\n")
cat("# of Journals in South Atlantic: ", count(EcologyBias$`Location of first author (region)` == "South Atlantic"), "\n")
cat("# of Journals in West North Central: ", count(EcologyBias$`Location of first author (region)` == "West North Central"), "\n")
cat("# of Journals in West South Central: ", count(EcologyBias$`Location of first author (region)` == "West South Central"), "\n")
cat("# of Journals in META ANALYSIS: ", count(EcologyBias$`Location of first author (region)` == "EMETA ANALYSIS"), "\n")
cat("# of Journals in Multiple: ", count(EcologyBias$`Location of first author (region)` == "Multiple"), "\n")
cat("# of Journals in Multiple Ranges: ", count(EcologyBias$`Location of first author (region)` == "Multiple Ranges"), "\n")

```


## Regions where the work was done (Popo defines)
```{r}
cat("# of Journals in East-North Central: ", count(EcologyBias$`Regions where the work was done  (Popo defines)` == "East-North Central"), "\n")
cat("# of Journals in Mid-Atlantic: ", count(EcologyBias$`Regions where the work was done  (Popo defines)` == "Mid-Atlantic"), "\n")
cat("# of Journals in Mid-South: ", count(EcologyBias$`Regions where the work was done  (Popo defines)` == "Mid-South"), "\n")
cat("# of Journals in Mountain: ", count(EcologyBias$`Regions where the work was done  (Popo defines)` == "Mountain"), "\n")
cat("# of Journals in New England: ", count(EcologyBias$`Regions where the work was done  (Popo defines)` == "New England"), "\n")
cat("# of Journals in North West: ", count(EcologyBias$`Regions where the work was done  (Popo defines)` == "North West"), "\n")
cat("# of Journals in Pacific: ", count(EcologyBias$`Regions where the work was done  (Popo defines)` == "Pacific"), "\n")
cat("# of Journals in South Atlantic: ", count(EcologyBias$`Regions where the work was done  (Popo defines)` == "South Atlantic"), "\n")
cat("# of Journals in West North Central: ", count(EcologyBias$`Regions where the work was done  (Popo defines)` == "West North Central"), "\n")
cat("# of Journals in West South Central: ", count(EcologyBias$`Regions where the work was done  (Popo defines)` == "West South Central"), "\n")
cat("# of Journals in META ANALYSIS: ", count(EcologyBias$`Regions where the work was done  (Popo defines)` == "EMETA ANALYSIS"), "\n")
cat("# of Journals in Multiple: ", count(EcologyBias$`Regions where the work was done  (Popo defines)` == "Multiple"), "\n")
cat("# of Journals in Multiple Ranges: ", count(EcologyBias$`Regions where the work was done  (Popo defines)` == "Multiple Ranges"), "\n")

```


```{r}

cat('# of Journal Entries in Coastal Ecosystems: ' , count(EcologyBias$`Ecosystem (Ped defines)` == "Coastal Ecosystems"), '\n')
cat('# of Journal Entries in Coniferous Forestss: ' , count(EcologyBias$`Ecosystem (Ped defines)` == "Coniferous Forests"), '\n')
cat('# of Journal Entries in Deciduous and Evergreens: ' , count(EcologyBias$`Ecosystem (Ped defines)` == "Deciduous and Evergreen"), '\n')
cat('# of Journal Entries in Ecosystems of the Deep Ocean: ' , count(EcologyBias$`Ecosystem (Ped defines)` == "Ecosystems of the Deep Ocean"), '\n')
cat('# of Journal Entries in Inland Aquatic Ecosystems: ' , count(EcologyBias$`Ecosystem (Ped defines)` == "Inland Aquatic Ecosystems"), '\n')
cat('# of Journal Entries in CLakes and Reservoirs Ecosystems: ' , count(EcologyBias$`Ecosystem (Ped defines)` == "Lakes and Reservoirs Ecosystems"), '\n')
cat('# of Journal Entries in Managed Terrestrial Ecosystems: ' , count(EcologyBias$`Ecosystem (Ped defines)` == "Managed Terrestrial Ecosystems"), '\n')
cat('# of Journal Entries in Polar & Alpine Tundra: ' , count(EcologyBias$`Ecosystem (Ped defines)` == "Polar & Alpine Tundra"), '\n')
cat('# of Journal Entries in Shrublands: ' , count(EcologyBias$`Ecosystem (Ped defines)` == "Shrublands"), '\n')
cat('# of Journal Entries in Temperate Deciduous Forests: ' , count(EcologyBias$`Ecosystem (Ped defines)` == "Temperate Deciduous Forests"), '\n')
cat('# of Journal Entries in Tropical Forests: ' , count(EcologyBias$`Ecosystem (Ped defines)` == "Tropical Forests"), '\n')
cat('# of Journal Entries in CWetland Forests: ' , count(EcologyBias$`Ecosystem (Ped defines)` == "Wetland Forests"), '\n')
cat('# of Journal Entries in Coniferous and Hardwood: ' , count(EcologyBias$`Ecosystem (Ped defines)` == "Coniferous and Hardwood"), '\n')
cat('# of Journal Entries in Coral Reefs: ' , count(EcologyBias$`Ecosystem (Ped defines)` == "Coral Reefs"), '\n')
cat('# of Journal Entries in Ecosystems of the Continental Shelves: ' , count(EcologyBias$`Ecosystem (Ped defines)` == "Ecosystems of the Continental Shelves"), '\n')
cat('# of Journal Entries in Estuaries and Enclosed Seas: ' , count(EcologyBias$`Ecosystem (Ped defines)` == "Estuaries and Enclosed Seas"), '\n')
cat('# of Journal Entries in Intertidal and Littoral Ecosystems: ' , count(EcologyBias$`Ecosystem (Ped defines)` == "Intertidal and Littoral Ecosystems"), '\n')
cat('# of Journal Entries in Managed Aquatic Ecosystems: ' , count(EcologyBias$`Ecosystem (Ped defines)` == "Managed Aquatic Ecosystems"), '\n')
cat('# of Journal Entries in Natural Grasslands: ' , count(EcologyBias$`Ecosystem (Ped defines)` == "Natural Grasslands"), '\n')
cat('# of Journal Entries in River and Stream Ecosystem: ' , count(EcologyBias$`Ecosystem (Ped defines)` == "River and Stream Ecosystem"), '\n')
cat('# of Journal Entries in Temperate Broad Leaved Evergreen: ' , count(EcologyBias$`Ecosystem (Ped defines)` == "Temperate Broad Leaved Evergreen"), '\n')
cat('# of Journal Entries in Temperate Deserts and Semi Deserts: ' , count(EcologyBias$`Ecosystem (Ped defines)` == "Temperate Deserts and Semi Deserts"), '\n')
cat('# of Journal Entries in Tropical Savannas: ' , count(EcologyBias$`Ecosystem (Ped defines)` == "CTropical Savannas"), '\n')
cat('# of Journal Entries in Multiple Ecosystems: ' , count(EcologyBias$`Ecosystem (Ped defines)` == "Multiple"), '\n')
cat('# of Journal Entries in META-ANALYSIS: ' , count(EcologyBias$`Ecosystem (Ped defines)` == "META-ANALYSIS"), '\n')
```


```{r}
cat('# of Journal Entries in Annelida: ' , count(EcologyBias$Phylum == "Annelida"), '\n')
cat('# of Journal Entries in Anthophyta: ' , count(EcologyBias$Phylum == "Anthophyta"), '\n')
cat('# of Journal Entries in Arthropoda: ' , count(EcologyBias$Phylum == "Arthropoda"), '\n')
cat('# of Journal Entries in Chordata: ' , count(EcologyBias$Phylum == "Chordata"), '\n')
cat('# of Journal Entries in Cnidaria: ' , count(EcologyBias$Phylum == "Cnidaria"), '\n')
cat('# of Journal Entries in Coniferophyta: ' , count(EcologyBias$Phylum == "Coniferophyta"), '\n')
cat('# of Journal Entries in Echinodermata: ' , count(EcologyBias$Phylum == "Echinodermata"), '\n')
cat('# of Journal Entries in Magnoliophyta: ' , count(EcologyBias$Phylum == "Magnoliophyta"), '\n')
cat('# of Journal Entries in Mollusca: ' , count(EcologyBias$Phylum == "Mollusca"), '\n')
cat('# of Journal Entries in Ochrophyta: ' , count(EcologyBias$Phylum == "Ochrophyta"), '\n')
cat('# of Journal Entries in Oomycota: ' , count(EcologyBias$Phylum == "Oomycota"), '\n')
cat('# of Journal Entries in Spermatophyta: ' , count(EcologyBias$Phylum == "Spermatophyta"), '\n')
cat('# of Journal Entries in Tracheophyta: ' , count(EcologyBias$Phylum == "Tracheophyta"), '\n')
cat('# of Journal Entries in Multiple: ' , count(EcologyBias$Phylum == "Multiple"), '\n')
```

```{r}
tally(EcologyBias$`Marine, Terrestrial, Freshwater, Mixed`)
```


```{r}
tally(EcologyBias$`Location of first author (region)`)
```

```{r}
tally(EcologyBias$`Regions where the work was done  (Popo defines)`)
```

```{r}
tally(EcologyBias$`Phylum`)
```

```{r}
chisq.test(tally(EcologyBias$`Marine, Terrestrial, Freshwater, Mixed`))

```
```{r}
chisq.test(tally(EcologyBias$`Regions where the work was done  (Popo defines)`))
```


```{r}
chisq.test(tally(EcologyBias$`Location of first author (region)`))
```

```{r}
chisq.test(tally(EcologyBias$`Phylum`))
```

```{r}
library(gplots)
# 1. convert the data as a table
dt <- as.table(as.matrix(tally(EcologyBias$`Regions where the work was done  (Popo defines)`)))
# 2. Graph
balloonplot(t(dt), main ="Based on Region where the work was done", xlab ="", ylab="",
            label = FALSE, show.margins = FALSE)
```


