---
title: "Dr.Daneshgar Data Analysis"
author: Veronica Marquez
output: html_notebook
---

get file location
```{r}
getwd()
```
run library
```{r}
library(mosaic)
```

```{r}
library(readxl)
EcologyBias_Daneshgar_Popo_2_ <- read_excel("/Users/veronicamarquez/Desktop/Dr.D Ecology Bias SC/MA321-Daneshgar/EcologyBias.Daneshgar.Popo.xlsx")
View(EcologyBias_Daneshgar_Popo_2_)
```

```{r}
EcologyBias <- EcologyBias_Daneshgar_Popo_2_[1:175, ]
```

Looking to see how many of each ecosystem exists in the data set 
```{r}
cat("# of Journals in Marine: ", count(EcologyBias$`Marine, Terrestrial, Freshwater, Mixed` == "Marine"), "\n")
cat("# of Journals in Terrestrial: ", count(EcologyBias$`Marine, Terrestrial, Freshwater, Mixed` == "Terrestrial"), "\n")
cat("# of Journals in Frshwater: ", count(EcologyBias$`Marine, Terrestrial, Freshwater, Mixed` == "Freshwater"), "\n")
cat("# of Journals in Mixed: ", count(EcologyBias$`Marine, Terrestrial, Freshwater, Mixed` == "Mixed"), "\n")
```
Notice here that there is a significant difference in the amount of articles that contain each ecosystem. Is this a glimpse into the "Sexy" type of ecosystem? Should I take it a step closer and analyze the ecosystems based on specif regions? If we take just this information, then we definitely some form of bias between these ecosystem in itself. 

##Location of First Author (Region)
```{r}
cat("# of Journals in East-North Central: ", count(EcologyBias$`Location of first author (region)` == "East-North Central"), "\n")
cat("# of Journals in Mid-Atlantic: ", count(EcologyBias$`Location of first author (region)` == "Mid-Atlantic"), "\n")
cat("# of Journals in Mid-South: ", count(EcologyBias$`Location of first author (region)` == "Mid-South"), "\n")
cat("# of Journals in Mountain: ", count(EcologyBias$`Location of first author (region)` == "Mountain"), "\n")
cat("# of Journals in New England: ", count(EcologyBias$`Location of first author (region)` == "New England"), "\n")
cat("# of Journals in North West: ", count(EcologyBias$`Location of first author (region)` == "North West"), "\n")
cat("# of Journals in Pacific: ", count(EcologyBias$`Location of first author (region)` == "Pacific"), "\n")
cat("# of Journals in South Atlantic: ", count(EcologyBias$`Location of first author (region)` == "South Atlantic"), "\n")
cat("# of Journals in West North Central: ", count(EcologyBias$`Location of first author (region)` == "West North Central"), "\n")
cat("# of Journals in West South Central: ", count(EcologyBias$`Location of first author (region)` == "West South Central"), "\n")
cat("# of Journals in META ANALYSIS: ", count(EcologyBias$`Location of first author (region)` == "EMETA ANALYSIS"), "\n")
cat("# of Journals in Multiple: ", count(EcologyBias$`Location of first author (region)` == "Multiple"), "\n")
cat("# of Journals in Multiple Ranges: ", count(EcologyBias$`Location of first author (region)` == "Multiple Ranges"), "\n")

```


## Regions where the work was done (Popo defines)
```{r}
cat("# of Journals in East-North Central: ", count(EcologyBias$`Regions where the work was done  (Popo defines)` == "East-North Central"), "\n")
cat("# of Journals in Mid-Atlantic: ", count(EcologyBias$`Regions where the work was done  (Popo defines)` == "Mid-Atlantic"), "\n")
cat("# of Journals in Mid-South: ", count(EcologyBias$`Regions where the work was done  (Popo defines)` == "Mid-South"), "\n")
cat("# of Journals in Mountain: ", count(EcologyBias$`Regions where the work was done  (Popo defines)` == "Mountain"), "\n")
cat("# of Journals in New England: ", count(EcologyBias$`Regions where the work was done  (Popo defines)` == "New England"), "\n")
cat("# of Journals in North West: ", count(EcologyBias$`Regions where the work was done  (Popo defines)` == "North West"), "\n")
cat("# of Journals in Pacific: ", count(EcologyBias$`Regions where the work was done  (Popo defines)` == "Pacific"), "\n")
cat("# of Journals in South Atlantic: ", count(EcologyBias$`Regions where the work was done  (Popo defines)` == "South Atlantic"), "\n")
cat("# of Journals in West North Central: ", count(EcologyBias$`Regions where the work was done  (Popo defines)` == "West North Central"), "\n")
cat("# of Journals in West South Central: ", count(EcologyBias$`Regions where the work was done  (Popo defines)` == "West South Central"), "\n")
cat("# of Journals in META ANALYSIS: ", count(EcologyBias$`Regions where the work was done  (Popo defines)` == "EMETA ANALYSIS"), "\n")
cat("# of Journals in Multiple: ", count(EcologyBias$`Regions where the work was done  (Popo defines)` == "Multiple"), "\n")
cat("# of Journals in Multiple Ranges: ", count(EcologyBias$`Regions where the work was done  (Popo defines)` == "Multiple Ranges"), "\n")

```


```{r}

cat('# of Journal Entries in Coastal Ecosystems: ' , count(EcologyBias$`Ecosystem (Ped defines)` == "Coastal Ecosystems"), '\n')
cat('# of Journal Entries in Coniferous Forestss: ' , count(EcologyBias$`Ecosystem (Ped defines)` == "Coniferous Forests"), '\n')
cat('# of Journal Entries in Deciduous and Evergreens: ' , count(EcologyBias$`Ecosystem (Ped defines)` == "Deciduous and Evergreen"), '\n')
cat('# of Journal Entries in Ecosystems of the Deep Ocean: ' , count(EcologyBias$`Ecosystem (Ped defines)` == "Ecosystems of the Deep Ocean"), '\n')
cat('# of Journal Entries in Inland Aquatic Ecosystems: ' , count(EcologyBias$`Ecosystem (Ped defines)` == "Inland Aquatic Ecosystems"), '\n')
cat('# of Journal Entries in CLakes and Reservoirs Ecosystems: ' , count(EcologyBias$`Ecosystem (Ped defines)` == "Lakes and Reservoirs Ecosystems"), '\n')
cat('# of Journal Entries in Managed Terrestrial Ecosystems: ' , count(EcologyBias$`Ecosystem (Ped defines)` == "Managed Terrestrial Ecosystems"), '\n')
cat('# of Journal Entries in Polar & Alpine Tundra: ' , count(EcologyBias$`Ecosystem (Ped defines)` == "Polar & Alpine Tundra"), '\n')
cat('# of Journal Entries in Shrublands: ' , count(EcologyBias$`Ecosystem (Ped defines)` == "Shrublands"), '\n')
cat('# of Journal Entries in Temperate Deciduous Forests: ' , count(EcologyBias$`Ecosystem (Ped defines)` == "Temperate Deciduous Forests"), '\n')
cat('# of Journal Entries in Tropical Forests: ' , count(EcologyBias$`Ecosystem (Ped defines)` == "Tropical Forests"), '\n')
cat('# of Journal Entries in CWetland Forests: ' , count(EcologyBias$`Ecosystem (Ped defines)` == "Wetland Forests"), '\n')
cat('# of Journal Entries in Coniferous and Hardwood: ' , count(EcologyBias$`Ecosystem (Ped defines)` == "Coniferous and Hardwood"), '\n')
cat('# of Journal Entries in Coral Reefs: ' , count(EcologyBias$`Ecosystem (Ped defines)` == "Coral Reefs"), '\n')
cat('# of Journal Entries in Ecosystems of the Continental Shelves: ' , count(EcologyBias$`Ecosystem (Ped defines)` == "Ecosystems of the Continental Shelves"), '\n')
cat('# of Journal Entries in Estuaries and Enclosed Seas: ' , count(EcologyBias$`Ecosystem (Ped defines)` == "Estuaries and Enclosed Seas"), '\n')
cat('# of Journal Entries in Intertidal and Littoral Ecosystems: ' , count(EcologyBias$`Ecosystem (Ped defines)` == "Intertidal and Littoral Ecosystems"), '\n')
cat('# of Journal Entries in Managed Aquatic Ecosystems: ' , count(EcologyBias$`Ecosystem (Ped defines)` == "Managed Aquatic Ecosystems"), '\n')
cat('# of Journal Entries in Natural Grasslands: ' , count(EcologyBias$`Ecosystem (Ped defines)` == "Natural Grasslands"), '\n')
cat('# of Journal Entries in River and Stream Ecosystem: ' , count(EcologyBias$`Ecosystem (Ped defines)` == "River and Stream Ecosystem"), '\n')
cat('# of Journal Entries in Temperate Broad Leaved Evergreen: ' , count(EcologyBias$`Ecosystem (Ped defines)` == "Temperate Broad Leaved Evergreen"), '\n')
cat('# of Journal Entries in Temperate Deserts and Semi Deserts: ' , count(EcologyBias$`Ecosystem (Ped defines)` == "Temperate Deserts and Semi Deserts"), '\n')
cat('# of Journal Entries in Tropical Savannas: ' , count(EcologyBias$`Ecosystem (Ped defines)` == "CTropical Savannas"), '\n')
cat('# of Journal Entries in Multiple Ecosystems: ' , count(EcologyBias$`Ecosystem (Ped defines)` == "Multiple"), '\n')
cat('# of Journal Entries in META-ANALYSIS: ' , count(EcologyBias$`Ecosystem (Ped defines)` == "META-ANALYSIS"), '\n')
```


```{r}
cat('# of Journal Entries in Annelida: ' , count(EcologyBias$Phylum == "Annelida"), '\n')
cat('# of Journal Entries in Anthophyta: ' , count(EcologyBias$Phylum == "Anthophyta"), '\n')
cat('# of Journal Entries in Arthropoda: ' , count(EcologyBias$Phylum == "Arthropoda"), '\n')
cat('# of Journal Entries in Chordata: ' , count(EcologyBias$Phylum == "Chordata"), '\n')
cat('# of Journal Entries in Cnidaria: ' , count(EcologyBias$Phylum == "Cnidaria"), '\n')
cat('# of Journal Entries in Coniferophyta: ' , count(EcologyBias$Phylum == "Coniferophyta"), '\n')
cat('# of Journal Entries in Echinodermata: ' , count(EcologyBias$Phylum == "Echinodermata"), '\n')
cat('# of Journal Entries in Magnoliophyta: ' , count(EcologyBias$Phylum == "Magnoliophyta"), '\n')
cat('# of Journal Entries in Mollusca: ' , count(EcologyBias$Phylum == "Mollusca"), '\n')
cat('# of Journal Entries in Ochrophyta: ' , count(EcologyBias$Phylum == "Ochrophyta"), '\n')
cat('# of Journal Entries in Oomycota: ' , count(EcologyBias$Phylum == "Oomycota"), '\n')
cat('# of Journal Entries in Spermatophyta: ' , count(EcologyBias$Phylum == "Spermatophyta"), '\n')
cat('# of Journal Entries in Tracheophyta: ' , count(EcologyBias$Phylum == "Tracheophyta"), '\n')
cat('# of Journal Entries in Multiple: ' , count(EcologyBias$Phylum == "Multiple"), '\n')
```

```{r}
tally(EcologyBias$`Marine, Terrestrial, Freshwater, Mixed`)
```


```{r}
tally(EcologyBias$`Location of first author (region)`)
```

```{r}
tally(EcologyBias$`Regions where the work was done  (Popo defines)`)
```

```{r}
tally(EcologyBias$`Phylum`)
```

```{r}
chisq.test(tally(EcologyBias$`Marine, Terrestrial, Freshwater, Mixed`))

```
```{r}
chisq.test(tally(EcologyBias$`Regions where the work was done  (Popo defines)`))
```


```{r}
chisq.test(tally(EcologyBias$`Location of first author (region)`))
```

```{r}
chisq.test(tally(EcologyBias$`Phylum`))
```

```{r}
library(gplots)
# 1. convert the data as a table
dt <- as.table(as.matrix(tally(EcologyBias$`Regions where the work was done  (Popo defines)`)))
# 2. Graph
balloonplot(t(dt), main ="Based on Region where the work was done", xlab ="", ylab="",
            label = FALSE, show.margins = FALSE)
```

```{r}
favstats(~(tally(EcologyBias$`Regions where the work was done  (Popo defines)`)), data = EcologyBias)
```
```{r}
favstats(~(tally(EcologyBias$`Phylum`)), data = EcologyBias)
```

```{r}
favstats(~(tally(EcologyBias$`Location of first author (region)`)), data = EcologyBias)
```

```{r}
Ecosystem <- EcologyBias[,16]
Ecosystem

class.clean <- na.omit(Ecosystem)
class.clean
```



```{r}
continent.work <- as.vector(EcologyBias$`Continent where work done`)
remove <- c ("Europe & Oceania", "META ANALYSIS","META ANALYSIS!")
continent.work %in% remove
continent.work <- continent.work [! continent.work %in% remove]
continent.work[43] <- "Oceania"
continent.work[61] <- "Oceania"
continent.work <- c(continent.work[1:97], continent.work[99:168])
continent.work
tally(continent.work)
chisq.test(tally(continent.work))

```
##State where work was done
```{r}
State.work <- as.vector(EcologyBias$`State or Municipality where the work was done`)
remove <- c ("Multiple", "META ANALYSIS","Multiple States")
State.work %in% remove
State.work <- State.work [! State.work %in% remove]

State.work <- State.work[!is.na(State.work)]
State.work
tally(State.work)
chisq.test(tally(State.work))
```
Our pvalue is still coming back the same 

##State where work was done, DELETED THE MULTIPLE STATE ONES WITHN GIVEN STATES
```{r}
State.work <- as.vector(EcologyBias$`State or Municipality where the work was done`)
remove <- c ("Multiple", "META ANALYSIS","Multiple States", "GA/FL", "OR/CA", "NY, MA, CT")
State.work %in% remove
State.work <- State.work [! State.work %in% remove]

State.work <- State.work[!is.na(State.work)]
State.work
tally(State.work)
chisq.test(tally(State.work))
```
our pvalue is still coming back the same 

```{r}
SquMil <- Square_Mileage[1:50, ]
SquMil
```
Fav stats for total Sq.Mil given all 50 states
```{r}
favstats(~(SquMil$'Total Sq. mi (mi^2)'), data = SquMil)
```

```{r}

```

fav stats for total Sq.Mil given Continent
```{r}
favstats(~(SquMil$'Sq. mi (million mi^2)'), data = SquMil)
```

```{r}
chisq.test(SquMil$`Total Sq. mi (mi^2)`)
```

# Chisq test ecosystems popo def
```{r}
eco.po <- as.vector(EcologyBias$`Marine, Terrestrial, Freshwater, Mixed`)
eco.po
remove <- c ("Mixed")
eco.po %in% remove
eco.po <- eco.po [! eco.po %in% remove]
eco.po

tally(eco.po)
chisq.test(tally(eco.po))
```

# Chisq ecosystem ped def
```{r}
eco.ped <- as.vector(EcologyBias$`Ecosystem (Ped defines)`)

remove <- c ("Multiple", "META-ANALYSIS")
eco.ped %in% remove
eco.ped <- eco.ped [! eco.ped %in% remove]

eco.ped <- c(eco.ped[1:50], eco.ped[52:169])
eco.ped
tally(eco.ped)
chisq.test(tally(eco.ped))
```
Continent where work was done cleaned 
```{r}
Continent.work <- as.vector(EcologyBias$`Continent where work done`)
remove <- c ("Europe & Oceania", "META ANALYSIS","META ANALYSIS!")
Continent.work %in% remove
Continent.work <- Continent.work [! Continent.work %in% remove]
Continent.work[43] <- "Oceania"
Continent.work[61] <- "Oceania"
Continent.work <- c(Continent.work[1:97], Continent.work[99:168])
Continent.work
tally(Continent.work)
chisq.test(tally(Continent.work))
```

# continent.work
```{r}
xchisq.test(tally(Continent.work))

```

# eco.po
```{r}
x <- tally(eco.po)
x <- as.matrix(x)
x
xchisq.test(x)

```
This data shows us the expected number each ecosystem is suppose to have given that all ecosystems had a fair chance of being "picked". This information explains why a pvalue soooooo small happened because comparing to the expected number and the actual number there is no close similarities. Which leads our findings to having bias in ecosystems.
# eco.ped
```{r}
xchisq.test(tally(eco.ped))
```

# Pie Chart of Ecosystem of Observed 
```{r}
slices <- c(20,127,12)
lbls <- c("Marine", "Terrestrial", "Freshwater")
pie(slices, labels = lbls, main="Pie Chart of Ecosystem (Observed)")
```

# Pie Chart of Ecosystem of Expected
```{r}
slicesExp <- c(53,53,53)
lbls <- c("Marine", "Terrestrial", "Freshwater")
pie(slicesExp, labels = lbls, main = "Pie Chart of Ecosystem (Expected)" )
```



```{r}
slices <- c(5,14,2,34,94,5,13)
lbls <- c("Africa", "Asia", "Australia", "Europe", " North America", "Oceania", "South America")
pie(slices, labels = lbls, main="Pie Chart of Continents where work was done")
```


```{r}
slices <- c(11.73,17.21,2.97,3.931,9.54,3.292,6.888)
lbls <- c("Africa", "Asia", "Australia", "Europe", " North America", "Oceania", "South America")
pie(slices, labels = lbls, main="Pie Chart of Sq. Mileage based on Continent")
```

# Chisq test Continent
```{r}
continent.work <- as.vector(EcologyBias$`Continent where work done`)
remove <- c ("Europe & Oceania", "META ANALYSIS","META ANALYSIS!")
continent.work %in% remove
continent.work <- continent.work [! continent.work %in% remove]
continent.work[43] <- "Oceania"
continent.work[61] <- "Oceania"
continent.work <- c(continent.work[1:97], continent.work[99:168])
continent.work
tally(continent.work)
chisq.test(tally(continent.work))

```
# Continent with Sqr mileage
```{r}
sm.earth <- 197 # sqr miles of earths surface

sm.continent <- c(SqrMileage$`Sq. mi (million mi^2)`)

sm.continent <- as.numeric(sm.continent[2:8])

prop.cont <- sm.continent/sum(sm.continent)
prop.cont
chisq.test(tally(continent.work), p=prop.cont)
xchisq.test(tally(continent.work),p=prop.cont )

```

# Chart for Continent with Sqr Mileage

```{r}
library(nhstplot)
plotchisqtest(254.462, 6)
```
# Chart using the Expected values 
```{r}
E <- c(35.26, 51.73, 8.93, 11.82, 28.67 ,9.89 ,20.70)
CS <- c("Africa", "Asia", "Aust", "Europe", "NA", "Oceania", "SA")
barplot(E, names.arg = CS, xlab ="Country", ylab = " # of Journals", col = "blue", main = "Expected # of Journals in each Continent based on the Sq. Mileage", border = "black")
```


# Chart using the Observed values
```{r}
O <- c(5,14,2,34,94,5,13)
C <- c("Africa", "Asia", "Aust", "Europe", "NA", "Oceania", "SA")
barplot(O, names.arg = C, xlab ="Country", ylab = " # of Journals", col = "blue", main = "Observed # of Journals in each Country based on Sqr. Mileage", border = "black")
```

# Combing both charts
```{r}

```


# Chisq test country
```{r}
country.work <- as.vector(EcologyBias$`Country where the work is done`)
remove <- c ("Brazil/Costa Rica/Puerto Rico", "Germany & New Zealand","Lesser Antilles/Guadelupe", "META ANALYSIS", "Multiple???", "US and Mexico", "Trinidad and Tobago")
country.work %in% remove
country.work <- country.work [! country.work %in% remove]
country.work[1:5] <- "USA"
country.work[8:11] <- "USA"
country.work <- c(country.work[1:25], country.work[27:34], country.work[36:77], country.work[79:95], country.work[97:121], country.work[123:144], country.work[146:168])
country.work
tally(country.work)
chisq.test(tally(country.work))

```
# Chisq test region
```{r}
region.work <- as.vector(EcologyBias$`Regions where the work was done  (Popo defines)`)
region.work
remove <- c ("Multiple", "META ANALYSIS", "Multiple Ranges", "MidSouth")
region.work %in% remove
region.work <- region.work [! region.work %in% remove]
region.work
region.work <- region.work[!is.na(region.work)]
region.work
tally(region.work)
chisq.test(tally(region.work))

```
# xChisq test Region
```{r}
tally(region.work)
xchisq.test(tally(region.work))
```

# Pie Chart of Region (Expected)
```{r}
Region.Obs <- c(6.70,6.70,6.70,6.70,6.70,6.70,6.70, 6.70, 6.70, 6.70)
R.names <- c("East.NC", "Mid-A", "Mid-S", "Mount", "New Eng", " NW", "Pacific", "S.Atl", "W.NC", "W.SC")
pie(Region.Obs, labels = R.names, main="Pie Chart of Region (Expected)")
```

# Pie Chart of Region (Observed)
```{r}
Region.Obs <- c(7,3,1,12,3,1,24, 10, 3, 3)
R.names <- c("East.NC", "Mid-A", "Mid-S", "Mount", "New Eng", " NW", "Pacific", "S.Atl", "W.NC", "W.SC")
pie(Region.Obs, labels = R.names, main="Pie Chart of Region (Observed)")
```


# Region with sqr mileage
```{r}
tally(region.work)
sm.region <- c(SqrMileage$`Sq. mi2 (mi^2)`)
sm.region[1:9]

sm.region <- as.numeric(sm.region[1:9])
prop.region <- sm.region/sum(sm.region)
prop.region
xchisq.test(tally(region.work),p=prop.region )
```
# Chart of Region based on Sqr Mileage Expected *Note no Mid-South
```{r}
RegionSM.Exp <- c(5.54,3.54,15.85,1.33,0.10,16.58,5.42, 9.41, 8.23)
Region <- c("East.NC", "Mid-A",  "Mount", "New Eng", " NW", "Pacific", "S.Atl", "W.NC", "W.SC")
barPlot1 <- barplot(RegionSM.Exp, names.arg = Region, xlab ="Region", ylab = " # of Journals", col = "dark green", main = "Expected # of Journals in each Region based on Sqr. Mileage", border = "black")
```


# Chart of Region based on Sqr Mileage Observed *Note no Mid-South
```{r}
RegionSM.Obs <- c(7,1,12,3,1,24, 10, 3, 3)
Region <- c("East.NC", "Mid-A",  "Mount", "New Eng", " NW", "Pacific", "S.Atl", "W.NC", "W.SC")
barPlot1 <- barplot(RegionSM.Obs, names.arg = Region, xlab ="Region", ylab = " # of Journals", col = "dark green", main = "Observed # of Journals in each Region based on Sqr. Mileage", border = "black")
```


# Chisq test state
```{r}
state.work <- as.vector(EcologyBias$`State or Municipality where the work was done`)
state.work
remove <- c ("Multiple States", "Multiple", "META ANALYSIS", "OR/CA", "GA/FL", "NY, MA, CT")
state.work %in% remove
state.work <- state.work [! state.work %in% remove]
state.work
state.work <- state.work[!is.na(state.work)]
state.work

tally(state.work)
xchisq.test(tally(state.work))

```
# Chart of State (Expected)
```{r}
state.EXP <- c(2.26,2.26,2.26,2.26,2.26,2.26,2.26,2.26,2.26,2.26,2.26,2.26,2.26,2.26,2.26,2.26,2.26,2.26,2.26,2.26,2.26,2.26,2.26,2.26,2.26,2.26,2.26)
state <- c("AK", "AL", "AZ", "CA", "CO", "FL", "HI", "IN", "KS", "MA", "MD", "MI", "MO", "MT", "NC", "NH", "NJ", "NM", "NY", "OH", "OR", "PR", "TX", "UT", "WA", "WI", "WY")
barPlot1 <- barplot(state.EXP, names.arg = state, xlab ="State", ylab = " # of Journals", col = "light green", main = "Expected # of Journals in each State", border = "black")
```

# Chart of State (Observed)
```{r}
state.obs <- c(2,2,3,13,2,4,1,1,2,1,1,4,1,1,1,1,2,1,1,1,4,2,3,1,3,1,2)
state <- c("AK", "AL", "AZ", "CA", "CO", "FL", "HI", "IN", "KS", "MA", "MD", "MI", "MO", "MT", "NC", "NH", "NJ", "NM", "NY", "OH", "OR", "PR", "TX", "UT", "WA", "WI", "WY")
barPlot1 <- barplot(state.obs, names.arg = state, xlab ="State", ylab = " # of Journals", col = "light green", main = "Observed # of Journals in each State ", border = "black")
```





# state with sqr mileage
```{r}
sm.state <- c(SqrMileage$`totalSQM`)
sm.state <- c(sm.state[1:27])
#sm.state
prop.state <- sm.state/sum(sm.state)
#prop.state
xchisq.test(tally(state.work),p=prop.state  )
tally(state.work)
```
# Chart of State with Sqr Mileage (Observed)
```{r}
state.obs <- c(2,2,3,13,2,4,1,1,2,1,1,4,1,1,1,1,2,1,1,1,4,2,3,1,3,1,2)
state <- c("AK", "AL", "AZ", "CA", "CO", "FL", "HI", "IN", "KS", "MA", "MD", "MI", "MO", "MT", "NC", "NH", "NJ", "NM", "NY", "OH", "OR", "PR", "TX", "UT", "WA", "WI", "WY")
barPlot1 <- barplot(state.obs, names.arg = state, xlab ="State", ylab = " # of Journals", col = "light green", main = "Observed # of Journals in each State based on Sqr. Mileage", border = "black")
```
# Chart of State with Sqr. Mileage (Expected)
```{r}
state.exp <- c(15.290,1.221,2.656, 3.813, 2.425, 1.397, 0.150, 0.848, 1.917, 0.824, 0289, 2.255, 1.624, 3.425, 1.227, 0.218, 0.203, 2.832, 1.269, 1.044, 2.292, 0.082, 6.256, 1.978, 1.661, 1.526, 2.278)
state <- c("AK", "AL", "AZ", "CA", "CO", "FL", "HI", "IN", "KS", "MA", "MD", "MI", "MO", "MT", "NC", "NH", "NJ", "NM", "NY", "OH", "OR", "PR", "TX", "UT", "WA", "WI", "WY")
barPlot1 <- barplot(state.exp, names.arg = state, xlab ="State", ylab = " # of Journals", col = "light green", main = "Expected # of Journals in each State based on Sqr. Mileage", border = "black")
```


# Chisq test ecosystems popo def
```{r}
eco.po <- as.vector(EcologyBias$`Marine, Terrestrial, Freshwater, Mixed`)
eco.po
remove <- c ("Mixed")
eco.po %in% remove
eco.po <- eco.po [! eco.po %in% remove]
eco.po

tally(eco.po)

```

# Chisq Test ecosystem popo def
```{r}
xchisq.test(tally(eco.po))
```

# Chart with Observed Values 
```{r}
Obs <- c(12, 20, 127)
EcoObs <- c("Freshwater", "Marine", "Terrestrial")
barPlot2 <- barplot(Obs, names.arg = EcoObs, xlab ="Ecosystem", ylab = " # of Journals", col = "purple", main = "Observed # of Journals in each Ecosystem", border = "black")
```


#Chart with Expected Value 
*given all ecosystems have a fair chance of being selected
*not based on sqr mileage just based on the "n"
```{r}
Exp <- c(53,53,53)
Eco <- c("Freshwater", "Marine", "Terrestrial")
barPlot1 <- barplot(Exp, names.arg = Eco, xlab ="Ecosystem", ylab = " # of Journals", col = "purple", main = "Expected # of Journals in each Ecosystem", border = "black")
```

# Trying to remove the most positive residual in Ecosystem *CASE 1*
```{r}
eco.NEW.po <- as.vector(EcologyBias$`Marine, Terrestrial, Freshwater, Mixed`)
eco.NEW.po
remove <- c ("Mixed", "Terrestrial")
eco.NEW.po %in% remove
eco.NEW.po <- eco.NEW.po [! eco.NEW.po %in% remove]
eco.NEW.po

```

```{r}
tally(eco.NEW.po)
xchisq.test(tally(eco.NEW.po))
```

```{r}
slicesExp <- c(12,20)
lbls <- c("Freshwater", "Marine")
pie(slicesExp, labels = lbls, main="Pie Chart of Ecosystem (Expected)")
slicesOBS <- c(16, 16)
pie(slicesOBS, labels = lbls, main = "Pie Chart of Ecosystem (Observed)")
```



```{r}
EcoExpected = matrix(c(53,53,53))
EcoExpected
EcoObserved = matrix(c(12,20,127))
EcoObserved
```


# Combining both graphs
```{r}
##counts <- table(EcosystemExp_Obs$EcoExpected, EcosystemExp_Obs$EcoObserved)
##barplot(counts, main = "Ecosystem", xlab = "# of Journals", col = c("purple", "Green"), legend = rownames(counts), beside = ##TRUE)
#require(gridExtra)
#grid.arrange(barPlot1, barPlot2 ,ncol = 2 )

```



# xChisq ecosystem po def with Sqr. Mileage
```{r}
sm.ecosystem <- c(Eco_Sqr_Mileage$`Sqr Mil (mil mi^2)`)

sm.ecosystem <- as.numeric(sm.ecosystem[1:3])

prop.cont <- sm.ecosystem/sum(sm.ecosystem)
prop.cont
chisq.test(tally(eco.po), p=prop.cont)
xchisq.test(tally(eco.po), p = prop.cont)
xchisq.test(tally(eco.po),p=prop.cont, simulate.p.value = TRUE )
```

# Pie Chart of Ecosystem with Sqr. Mileage Expected
```{r}
slicesExp <- c(110.74,44.92,3.34)
lbls <- c("Marine", "Terrestrial", "Freshwater")
pie(slicesExp, labels = lbls, main="Pie Chart of Ecosystem (Expected)")
```
# Chart of Ecosystem with Sqr. Mileage Expected
```{r}
Exp <- c(3.34, 110.74, 44.92)
EcoExp <- c("Freshwater", "Marine", "Terrestrial")
barPlot2 <- barplot(Exp, names.arg = EcoExp, xlab ="Ecosystem", ylab = " # of Journals", col = "green", main = "Expected # of Journals in each Ecosystem based on Sqr. Mileage", border = "black")
```




# Pie Chart of Ecosytem with Sqr. Mileage Observed
```{r}
slices <- c(20,127,12)
lbls <- c("Marine", "Terrestrial", "Freshwater")
pie(slices, labels = lbls, main="Pie Chart of Ecosystem (Observed)")
```
# Chart of Ecosystem with Sqr Mileage Observed
```{r}
Obs <- c(12, 20, 127)
EcoObs <- c("Freshwater", "Marine", "Terrestrial")
barPlot2 <- barplot(Obs, names.arg = EcoObs, xlab ="Ecosystem", ylab = " # of Journals", col = "green", main = "Observed # of Journals in each Ecosystem based on Sqr. Mileage", border = "black")
```




# Chisq ecosystem ped def
```{r}
eco.ped <- as.vector(EcologyBias$`Ecosystem (Ped defines)`)

remove <- c ("Multiple", "META-ANALYSIS")
eco.ped %in% remove
eco.ped <- eco.ped [! eco.ped %in% remove]

eco.ped <- c(eco.ped[1:50], eco.ped[52:169])
eco.ped
tally(eco.ped)
chisq.test(tally(eco.ped))
```

# Chisq Country where work was done
```{r}
tally(country.work)
chisq.test(tally(country.work))
```

# XChisq Country where work was done
```{r}
xchisq.test(tally(country.work))
```

# XChisq Country where work was done based on square mileage
```{r}
sm.country <- c(CountrySqrMileage$`Sqr.Mil (mi^2)`)

sm.country <- as.numeric(sm.country[1:37])

prop.cont <- sm.country/sum(sm.country)
prop.cont
chisq.test(tally(country.work), p=prop.cont)
xchisq.test(tally(country.work), p = prop.cont)
xchisq.test(tally(country.work),p=prop.cont, simulate.p.value = TRUE )
```


